---
title: "HuYu2_TimeSeries1"
author: "Zhaowei Cai"
date: 'Mar 26,2024 - Mar 29, 2024'
output:
  pdf_document: default
  html_document: default
  word_document: default
---


# Loading packages
```{r}
library(tseries)
library(forecast)
library(stringr)
library(PerformanceAnalytics)
library(xts)
```

# Part 1
```{r}
set.seed(123)
#create a time series with right observations and first element is 0
ts_sim <- arima.sim(list(order = c(1,1,0), ar=0.65), n = 635)

left <- 600
right <- 636
it <- left:right
```


```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, left, right), lty="dotted", col=4)
```

# Part 2


```{r}
myts = subset(ts_sim, subset=rep(c(TRUE, FALSE), times=c(600, 36)))
```

## Step 1: visualize myts
```{r}
plot.zoo(myts, col=4, las=1, main="Time Series")
```
## Step 2: unit root test (augmented Dickey-Fuller) of myts
```{r}
adf.test(myts, alternative = 'stationary')
```
P-value greater than 0.05, not reject H0, and it is not stationary.

## Step 3: differentiate myts, creating mydts
```{r}
mydts = diff(myts)
```

## Step 4: unit root test (augmented Dickey-Fuller) of mydts
```{r}
adf.test(mydts, alternative = 'stationary')
```
P-value less than 0.05, reject H0, and it is stationary.

## Step 5: identify lags for mydts
```{r}
par(mfrow=c(1,2), mar=c(5,4,3,3))
acf(mydts)
pacf(mydts)
```
ACF decreases slowly, but PACF shows that it is an AR(1) (lag=1 is relevant only).

## Step 6: train the model with auto.arima for mydts
```{r}
fit_mydts = auto.arima(
  mydts,
  max.p = 3,
  max.q = 3,
  ic = "aicc",
  seasonal = FALSE,
  stationary = TRUE,
  lambda = NULL,
  stepwise = FALSE,
  approximation = FALSE
)
summary(fit_mydts)
```

## Step 7: fit the original time series, i.e. myts

```{r}
fit_myts = arima(myts, c(1, 1, 0))
summary(fit_myts)
```

Or can directly fit the original time series
```{r}
fit_myts2 = auto.arima(myts)
summary(fit_myts2)
```

# Part 3

## Part 3(a)
```{r}
forecast_myts = forecast(fit_myts, h=36, level=0.95)
plot(forecast_myts, col=4, las=1)
abline(v=c(1, 600, 636), lty="dotted", col=4)
lines(601:636, ts_sim[601:636], lty="dotted", col="red")
# red is observation and blue is prediction
```

## Part 3(b)
```{r}
# since it is one step ahead predictin, so we need use for loop
pred_df <- data.frame(NULL)
for(t in 600:636){
  pred_onestep <- forecast(ts_sim[1:t], h=1, level=0.95, model = fit_myts)
  pred_df <- rbind(pred_df, data.frame(mean = pred_onestep$mean[1], lower = pred_onestep$lower[1], upper = pred_onestep$upper[1]))
}
```

```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, left, right), lty="dotted", col=4)

lines(it, pred_df$mean, col = 'red')
lines(it, pred_df$lower, col = 'green')
lines(it, pred_df$upper, col = 'green')
legend(40, 40, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```


# Part 4

Generate AR(1) model data
```{r}
set.seed(123)
ts_sim = arima.sim(list(ar=0.65),n=636)
```
add trend in the data
```{r}
ts_sim=ts_sim + 0.33 + 0.15*time(ts_sim)  
```
Generate plots
```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, 600, 636), lty="dotted", col=4)
```

# Part 5

```{r}
myts = subset(ts_sim, subset=rep(c(TRUE, FALSE), times=c(600, 36)))
```
Step 1: visualize myts
```{r}
plot.zoo(myts, col=4, las=1, main="Time Series")
```
Fit trend part
```{r}
time = time(myts)
reg=lm(myts~time)
summary(reg)
```

```{r}
plot(as.vector(time(myts)), rstudent(reg), ylab="Standardized Res.", xlab="Time", type="o")
```

```{r}
z=rstandard(reg)
par(mfrow=c(1,2))
acf(z)
pacf(z)
```
Remove the trend part and fit the residuals
```{r}
newts=ts(residuals(reg))
```

Step 6: train the model with auto.arima for newts

```{r}
fit_newts = auto.arima(newts, max.p=3, max.q=3, ic="aicc",
                       seasonal=FALSE, stationary=TRUE, lambda=NULL, 
                       stepwise=FALSE, approximation=FALSE
                       )
summary(fit_newts)
```

# Part 6

## Part 6(a)
```{r}
prediction <- forecast(fit_newts, h=length(it), level=0.95)
pred_df <- data.frame(time = it)
pred_df$mean <- prediction$mean + predict(reg, newdata = data.frame(time = it))
pred_df$lower <- prediction$lower + predict(reg, newdata = data.frame(time = it))
pred_df$upper <- prediction$upper + predict(reg, newdata = data.frame(time = it))
```

```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, left, right), lty="dotted", col=4)
lines(it, pred_df$mean, col = 'red')
lines(it, pred_df$lower, col = 'green')
lines(it, pred_df$upper, col = 'green')
legend(40, 150, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```
```{r}
plot(it, ts_sim[it], col=4, las=1, type = 'l')
abline(v=c(left, right), lty="dotted", col=4)
lines(it, pred_df$mean, col = 'red')
lines(it, pred_df$lower, col = 'green')
lines(it, pred_df$upper, col = 'green')
legend(left, 128, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```
## Part 6(b)
Predict the residuls part
```{r}
# since it is one step ahead predictin, so we need use for loop

# generate the residuals
rests <- ts_sim - predict(reg, newdata = data.frame(time = time(ts_sim)))

pred_df <- data.frame(NULL)
for(t in it){
  pred_onestep <- forecast(rests[1:t], h=1, level=0.95, model = fit_newts)
  pred_df <- rbind(pred_df, data.frame(mean = pred_onestep$mean[1], lower = pred_onestep$lower[1], upper = pred_onestep$upper[1]))
}
```

add predicted trend back
```{r}
pred_df$mean <- pred_df$mean + predict(reg, newdata = data.frame(time = it))
pred_df$lower <- pred_df$lower + predict(reg, newdata = data.frame(time = it))
pred_df$upper <- pred_df$upper + predict(reg, newdata = data.frame(time = it))
```

plot pred, obs and CI

```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, 600, 636), lty="dotted", col=4)
lines(it, pred_df$mean, col = 'red')
lines(it, pred_df$lower, col = 'green')
lines(it, pred_df$upper, col = 'green')
legend(40, 120, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```

```{r}
plot(it, ts_sim[it], col=4, las=1, type = 'l')
abline(v=c(600, 636), lty="dotted", col=4)
lines(it, pred_df$mean, col = 'red')
lines(it, pred_df$lower, col = 'green')
lines(it, pred_df$upper, col = 'green')
legend(left, 130, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```

## Part 7

```{r}
#(a)
data = read.csv('TSLA1.csv')

library(forecast)
library(zoo)
library(tseries)
TSLA = data$Close
time = as.Date(data$Date, format = '%m/%d/%y')
df = data.frame(datefield = time, TSLA = TSLA)
TSLA_stock = with(df, zoo(TSLA, order.by = time))
plot.zoo(TSLA_stock, col=4, las=1, main="TSLA")

# Use the closing price to get log return
log_return = na.omit(diff(log(data$Close))) # log return
time = as.Date(data$Date, format = '%m/%d/%y')[-1]
df = data.frame(datefield = time, TSLA = log_return)
TSLA_return = with(df, zoo(TSLA, order.by = time))
plot(df, main = "TSLA log returns")
plot.zoo(TSLA_return, col=4, las=1, main="TSLA")
# It seems there is no drift or a trend from the plot.
# We can use Augmented Dickey-Fuller Test Unit Root Test for more details:

library(urca)
summary(ur.df(log_return, type='trend', lags=20, selectlags="BIC"))
# no time trend, no drift

#(b)
tseries::adf.test(TSLA_return)
model <- auto.arima(TSLA_return, ic = 'bic',stationary = T, trace = T)
model # BIC = -6427.534 ARIMA(2,0,0)

# (c)
predict(model, newdata = data.frame(time = '2020-01-02'))
# The closing price of Jan 2, 2020:
exp(log(TSLA[1510])+TSLA_return[1509])
```